Kepler's Demographic Cliff: Ages of Young Stars and Planets in the Kepler Field

### 2024-05-14 paper/ms.tex
update ms

Exoplanet science is in an age of abundance, fueled by the discovery
of thousands of worlds orbiting close to their host stars
\citep{Borucki10,2015JATIS...1a4003R}.  Most of these exoplanets are
billions of years old.  This fact leaves many gaps in our knowledge of
the exact physical and dynamical pathways that produce these objects.
The reason is that processes such as thermal cooling
\citep{2007ApJ...659.1661F}, atmospheric loss
\citep{2019AREPS..47...67O}, giant impacts
\citep{2014prpl.conf..595R}, and dynamical instabilities
\citep{2017MNRAS.470.1750I} are expected to be most efficient over
$\approx$10\,Myr to $\approx$1\,Gyr timescales. For most known
exoplanets, these processes have already run their course.

@@

The main problem with the ``cluster-first'' approach is that most
nearby young stars are in the field.  Only $\approx$1\% of stars
within the nearest $\approx$500\,pc are associated with their birth
cluster
\citep[e.g.][]{Zari_2018,CantatGaudin_2020,Kounkel_2020,Kerr_2021}.
One can appreciate the problem by querying NASA's Exoplanet Archive
\citep[NEA;][]{2013PASP..125..989A} for transiting planets younger
than 1\,Gyr.  Imposing reasonable precision requirements
(e.g.~$t$$<$1\,Gyr at 2$\sigma$, or else ages precise to say S/N$=$3)
gives $\approx$50 ``young'' transiting planets at the time of writing.
Most of these planets are in clusters, and were found by Kepler, K2,
or TESS.  However, combined, those surveys have discovered
$\approx$5{,}000 planets.  Unless there is a major bias against the
discovery of sub-gigayear old planets with precisely constrained ages,
a constant star formation history implies that an order of magnitude
more young transiting planets should be known.

This study aims to resolve two main questions.  First, how wrong is it
to assume a uniform age distribution for transiting planet host stars?
Second, where are the missing young planets?  Briefly, we will find
that uniform is wrong by a factor of a few, and that at least one bias
against the discovery of young planets is in the measurement of their
ages.

@@

However, beyond rotation-based ages, recent years have also yielded
improvements in the lithium age scale.  Lithium ages include two
distinct regimes: depletion boundary ages for M-dwarfs in star
clusters, and the less precise decline-based ages for individual field
FGK stars \citep{Soderblom_2010}.  Here we focus on the latter
approach, which relies on the observation that the Li abundances of
partially-convective stars gradually decline as they age
\citep[e.g.][]{2005A&A...442..615S}.  The theoretical explanation for
this decline is debated
\citep[e.g.][]{1995ApJ...441..865C,2010ApJ...716.1269D,2019MNRAS.485.4052C}.
Empirical understanding has however improved due to the work by
\citet{Jeffries_2023}, who built a model for how the equivalent width
(EW) of the \ion{Li}{1} 6708\,\AA\ doublet evolves as a function of
stellar effective temperature and age for a set of 6{,}200 stars in 52
open clusters.  Two-sided lithium ages are useful for Kepler (FGK)
stars between $\approx$0.03-0.5\,Gyr, with a strong dependence on
spectral type since the K dwarfs lose their surface lithium much
faster than G dwarfs.  The precision of the ages reported by the
\citet{Jeffries_2023} method in this regime are in the range of
0.3-0.5~dex.

@@

This work is focused on Kepler stars for which ages can be inferred
using either rotation, lithium, or both.  These ``age-dateable stars''
are a minority of the \nkeplerstars\ Kepler targets.  Rotation periods
have been reported for $\approx$one in three Kepler targets
\citep[e.g.][]{McQuillan_2014,Santos_2021}.  High-resolution spectra
suitable for measurement of the \ion{Li}{1} 6708\,\AA\ doublet have
only been acquired for the Kepler objects of interest (KOIs), which
comprise a few percent of Kepler's targets.  In the following, we will
describe the set of stars that we adopt with measured rotation periods
(Section~\ref{subsec:rotsel}), the set of objects we adopt as planets
(Section~\ref{subsec:planetsel}), and the set of planets with
high-resolution spectra suitable for lithium analysis
(Section~\ref{subsec:lithiumsel}).  Figure~\ref{fig:stellarprops}
provides a visualization of the various samples.

@@

To select stars with rotation periods, we turn to previous work.  Many
investigators have derived rotation periods both specifically for KOIs
\citep{McQuillan_2013,Walkowicz_2013,Mazeh_2015,Angus_2018,David_2021},
and also for the broader set of all Kepler target stars
\citep{McQuillan_2014,Reinhold_2015,Santos_2019,Santos_2021}.  These
studies used a range of detection methods and selection functions.  We
are interested in understanding the age distribution of all Kepler
target stars irrespective of KOI status.  Out of these studies, the
only homogeneous analysis of both Kepler targets and KOIs were the
studies by \citet{Santos_2019,Santos_2021} (\citetalias{Santos_2019}
and \citetalias{Santos_2021}).  \citetalias{Santos_2019} and
\citetalias{Santos_2021} combined a wavelet analysis and
autocorrelation-based approach, and cumulatively reported rotation
periods for 55{,}232 main-sequence and subgiant FGKM stars.  They
included known KOIs and binaries in their analysis, and assigned them
specific quality flags, while removing transits and eclipses during
the stellar rotation measurement process.

@@

We considered planets in the NEA cumulative KOI table, which included
the best knowledge available on any given planet candidate as of 6
June 2023, while also incorporating human-based vetting.  These
planets represent a superset of those in the fully automated Q1-Q17
DR25 KOI Table \citep{Thompson_2018}, which could be adopted in future
work for planet occurrence rate calculations.  This version of the
cumulative KOI table included \nkoisnofp\ objects that are either
``confirmed'' or ``candidate'' planets, after excluding known false
positives.

@@

% stat from _estimate_new_lithium_timefrac.py
Lithium equivalent widths and abundances for the Kepler Objects of
Interest were already analyzed by \citet{2018ApJ...855..115B} for
$\approx$three quarters of the spectra in our sample.  However, new
spectra have since been acquired, and our approach and selection
function are different.  We therefore perform our own processing of
the reduced HIRES spectra.

We queried the CPS archive to collect all blaze-corrected HIRES
spectra that were available for all non-false positive Kepler Objects
of Interest with a MES$>$10.  This yielded at least one spectrum for
\nlithiumstars\ stars and \nlithiumplanets\ planets.  For stars with
multiple spectra available at varying ranges of signal-to-noise (S/N),
we analyzed only the spectrum with the highest number of counts.  The
resulting spectra were acquired between 6 September 2009 and 16 May
2024, and cumulatively comprise \nhireshours\,hours of open-shutter
time.

@@

Our EW measurement approach does not correct for the neighboring
\ion{Fe}{1} 6707.44\,\AA\ blend.  To evaluate the accuracy and
precision of our measurement technique, after applying an initial
iteration on the \nlithiumstars\ KOIs with spectra, we compared our
lithium equivalent widths with those reported by
\citet{2018ApJ...855..115B}.  For the \nbergeroverlap\ stars in both
samples, we found broad agreement at $>$30\,m\AA, and significant
differences at $<$30\,m\AA\ because \citet{2018ApJ...855..115B}
required positive EWs, while we allow for negative ones.  At
$>$30\,m\AA, there is however a small offset in our respective scales,
such that (B24-B18)$=$7.50$\pm$8.55\,m\AA.  This offset is caused by
our non-treatment of the iron blend.  We therefore directly subtract
the constant mean value noted above when calculating lithium-based
ages.

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.47\textwidth]{li_vs_teff_koi_X_JUMP_eagles.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Equivalent widths (EW) of the \ion{Li}{1} 6708\,\AA\ doublet
    for planet-hosting stars.} These measurements were made from
    Keck/HIRES spectra collected between 2009 and 2024 that comprised
    \nhireshours\,hours of open-shutter time.  Lines are the ``mean''
    isochrones from the same models as in Figure~\ref{fig:models}; the
    intrinsic dispersion beyond $\approx$1\,Gyr becomes much larger
    than changes in the mean.  Some stars with lithium detections do
    not have detected rotation periods, and vice-versa.
		\label{fig:li_vs_teff}
	}
\end{figure}

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{gyromodeldispersion.pdf}
		}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{li_vs_teff_koi_X_S19S21dquality_eagles_showpoints_nodata.pdf}
		}
	\end{center}
	\vspace{-0.6cm}
	\caption{
		{\bf The models.}
    Points represent $10^4$ draws from models that have been fitted to
    rotation periods \citep{Bouma_2023} and lithium equivalent widths
    \citep[EWs;][]{Jeffries_2023} of stars in open clusters.  Lines
    are the ``mean models'' at various ages; the intrinsic dispersion
    in rotation and lithium about these mean models, which is what the
    models fit, sets the theoretical precision floor for the
    age-dating methods.    Additional sources of uncertainty,
    including measurement uncertainty, impose further limits on
    achievable precision.  These models are calibrated using open
    clusters younger than 4\,Gyr.  The displayed points assume a
    uniform distribution in temperature for visual clarity.  The sizes
    of the points are the same in each panel, so that low apparent
    density signifies greater dispersion around the mean.
		\label{fig:models}
	}
\end{figure*}

@@

Using our effective temperatures and rotation periods, we ran
\texttt{gyro-interp} over an age grid of 0 to 4\,Gyr, linearly spaced
over 500 grid points.  \texttt{gyro-interp} itself is calibrated out
to 4\,Gyr, the age of M67
\citep[see][]{2022ApJ...938..118D,Gruner_2023}.  To estimate the
stellar spin-down rate as a function of time and stellar temperature,
the framework interpolates between reference open clusters using cubic
hermite polynomials.  We calculate the probability of the
rotation-based age $t_{\rm gyro}$, given the observed rotation
periods, effective temperatures, and their uncertainties by
integrating Equation~1 of \citet{Bouma_2023} at the default grid
resolution.  This procedure marginalizes over the intrinsic age and
temperature dependent scatter that is observed in the cluster
sequences to yield statistical age uncertainties.

@@

\subsection{Planet Quality Flags}
\label{subsec:plflags}
Some planets are more reliable than others.  We use the following
additional quality indicators to assess the reliability and utility of
a planet, and assemble them into a separate bit-flag, $Q_{\rm
planet}$.

@@

In the planet sample, the cumulative KOI table included
\nnonfpcumkois\ confirmed and candidate planets, orbiting
\nnonfpcumkoihosts\ stars.  Assume that we apply all the same stellar
cuts, save for that on RUWE.  Then, require these confirmed and
candidate planets to have MES$>$10.  This combination of selection
criteria yield \nplwgyroagewithgrazingandhighruwe\ planets with
rotation-based ages orbiting
\nplhoststarwgyroagewithgrazingandhighruwe\ stars
(\nplhoststarwgyroagejusthighruwe\ of these stars with ``high'' RUWE).
If we additionally require the planets to be non-grazing ($b<0.8$), we
are left with \nplwgyroagenograzing\ planets orbiting
\nplhoststarwgyroagenograzing\ stars.

@@

To calculate lithium ages, we use \texttt{EAGLES} (git commit
\texttt{ac09637}), which, similar to \texttt{gyro-interp}, is based on
a purely empirical interpolation approach.  \texttt{EAGLES} however
was calibrated on lithium measurements of stars observed by the
Gaia-ESO spectroscopic survey in 52 open clusters with ages spanning 2
to 6{,}000\,Myr \citep{Jeffries_2023}.  We adopt a linear prior in
age, spanning $10^6$ to $10^{10}$\,yr, and symmetrize our EW
uncertainties for purposes of interfacing with \texttt{EAGLES} by
taking the maximum of the measured positive and negative 1$\sigma$
uncertainty intervals.

@@

Figure~\ref{fig:hist_tgyro} shows the gyrochone age distributions of
stars and known planet-hosts in the Kepler field.  In detail, these
``histograms'' are sampled from the posterior probability
distributions by drawing ten random samples from each posterior
distribution, and then computing the histogram of the resulting
samples.  We truncated the plots near $4$\,Gyr since this is the
cluster age out to which our interpolation-based method is calibrated.
For Sun-like stars, the fraction of rotation signals detectable by
Kepler has been found to be near-unity up to $\approx$3\,Gyr, and it
drops to almost zero by $\approx$5\,Gyr \citep{2022ApJ...937...94M}.
This trend has the opposite functional form from our derived age
distributions.

@@

In terms of detected planet counts, our rotation-based ages for the
Kepler sample yield \{\nplyounggyro, \nplmidgyro, \nploldgyro\}
detected planets in the 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr bins.  For
the sub-gigayear bin in particular, which holds extra weight in
studies of planet evolution, while there are \nplyounggyro\ planets
with median ages below 1\,Gyr, requiring $t_{\rm gyro}$$<$1\,Gyr at
2$\sigma$ confidence yields \nplyounggyrotwosigma\ planets orbiting
\nplhostsyounggyrotwosigma\ stars.  The stellar and planetary
properties of this latter sample of young planets are given in
Table~\ref{tab:planets}.

@@

What are the most notable confirmed planets with well-measured ages?
These planets can be split into two classes: those in clusters, and
those with independent and consistent rotation and lithium-based ages.

@@

There are currently four known stellar ensembles in the Kepler field
that have yielded a total of fourteen transiting planets.  Our
analysis blindly recovers each of these planets, as well as their
ages.

@@

We compared the independent rotation and lithium-based ages using the
point-estimates of the posteriors, when available.  There are three
cases of interest:

@@

In the entire sample (i.e.~without imposing any quality cuts) this
yielded \allagesyesconsistent\ consistent cases,
\allagesmaybeconsistent\ ``maybe consistent'' cases, and
\allagesnoconsistent\ inconsistent cases.  Rephrased, in the overall
sample, \fracconsistentallages\% of stars have consistent rotation and
lithium-based ages (\fracpotentiallyconsistentallages\% potentially
consistent).

>>>>

Exoplanet science is enjoying an age of abundance, fueled by the discovery
of thousands of worlds orbiting close to their host stars
\citep{Borucki10,2015JATIS...1a4003R}.  Most of these exoplanets are
billions of years old.  This fact leaves many gaps in our knowledge of
the exact physical and dynamical pathways that produce these objects.
The reason is that processes such as thermal cooling
\citep{2007ApJ...659.1661F}, atmospheric loss
\citep{2019AREPS..47...67O}, giant impacts
\citep{2014prpl.conf..595R}, and dynamical instabilities
\citep{2017MNRAS.470.1750I} are expected to be most efficient over
$\approx$10\,Myr to $\approx$1\,Gyr timescales. For most known
exoplanets, these processes have already run their course.

@@

The main problem with the ``cluster-first'' approach is that most
nearby sub-gigayear stars are in the field.  In other words,
$\lesssim$1\% of stars within the nearest $\approx$500\,pc have been
associated with their birth cluster
\citep[e.g.][]{Zari_2018,CantatGaudin_2020,Kounkel_2020,Kerr_2021}.
One can appreciate the problem by querying NASA's Exoplanet Archive
\citep[NEA;][]{2013PASP..125..989A} for transiting planets younger
than 1\,Gyr.  Imposing reasonable precision requirements
(e.g.~$t$$<$1\,Gyr at 2$\sigma$, or else ages precise to say S/N$=$3)
gives $\approx$50 ``young'' transiting planets at the time of writing.
Most of these planets are in clusters, and were found by Kepler, K2,
or TESS.  However, combined, those surveys have discovered
$\approx$5{,}000 planets.  
Barring a major discovery bias against sub-gigayear planets,
a constant star formation history would imply that an order of magnitude
more young transiting planets should be known.

This study aims to resolve two main questions.  First, how wrong is it
to assume a uniform age distribution for transiting planet host stars?
Second, where are the missing young planets?  Briefly, we will find
that uniform is wrong by a factor of a few, and that increased stellar
activity for young stars may be a less important bias than the
challenge of determining precise ages for field stars.

@@

However, beyond rotation-based ages, recent years have also yielded
improvements in the lithium age scale.  Lithium ages include two
distinct regimes: depletion boundary ages for M dwarfs and brown dwarfs in star
clusters, and the less precise decline-based ages for individual field
FGK stars \citep{Soderblom_2010}.  Here we focus on the latter
approach, which relies on the observation that the Li abundances of
partially-convective stars gradually decline as they age
\citep[e.g.][]{2005A&A...442..615S}.  The theoretical explanation for
this decline is debated
\citep[e.g.][]{1995ApJ...441..865C,2010ApJ...716.1269D,2019MNRAS.485.4052C}.
Empirical understanding has however improved due to the work by
\citet{Jeffries_2023}, who built a model for how the equivalent width
(EW) of the \ion{Li}{1} 6708\,\AA\ doublet evolves as a function of
stellar effective temperature and age for a set of 6{,}200 stars in 52
open clusters.  Two-sided lithium ages are useful for Kepler (FGK)
stars between $\approx$0.03-0.5\,Gyr, with a strong dependence on
spectral type since the K dwarfs lose their surface lithium much
faster than G dwarfs.  The precision of the ages reported by the
\citet{Jeffries_2023} method in this regime are in the range of
0.3-0.5~dex.

@@

This work is focused on Kepler stars for which ages can be inferred
using either rotation, lithium, or both.  These ``age-dateable stars''
are a minority of the \nkeplerstars\ Kepler targets.  Rotation periods
have been reported for roughly one in three Kepler targets
\citep[e.g.][]{McQuillan_2014,Santos_2021}.  High-resolution spectra
suitable for measurement of the \ion{Li}{1} 6708\,\AA\ doublet have
only been acquired for the Kepler objects of interest (KOIs), which
comprise a few percent of Kepler's targets.  In the following, we will
describe the set of stars that we adopt with measured rotation periods
(Section~\ref{subsec:rotsel}), the set of objects we adopt as hosting planets
(Section~\ref{subsec:planetsel}), and the subset of these with
high-resolution spectra suitable for lithium analysis
(Section~\ref{subsec:lithiumsel}).  Figure~\ref{fig:stellarprops}
provides a visualization of the various samples.

@@

To select stars with rotation periods, we turn to previous work.  Many
investigators have derived rotation periods both specifically for KOIs
\citep{McQuillan_2013,Walkowicz_2013,Mazeh_2015,Angus_2018,David_2021},
and also for the broader set of all Kepler target stars
\citep{McQuillan_2014,Reinhold_2015,Santos_2019,Santos_2021}.  These
studies used a range of detection methods and selection functions.  We
are interested in understanding the age distribution of all Kepler
target stars irrespective of KOI status.  Out of these studies, the
only homogeneous analysis of both Kepler targets and KOIs were the
studies by \citet{Santos_2019,Santos_2021}, hereafter \citetalias{Santos_2019}
and \citetalias{Santos_2021}.  \citetalias{Santos_2019} and
\citetalias{Santos_2021} combined a wavelet analysis and
autocorrelation-based approach, and cumulatively reported rotation
periods for 55{,}232 main-sequence and subgiant FGKM stars.  They
included known KOIs and binaries in their analysis, and assigned them
specific quality flags, while removing transits and eclipses during
the stellar rotation measurement process.

@@

We considered planets in the NEA cumulative KOI table as of 6 June
2023, which included the best knowledge available on any given planet
candidate while also incorporating human-based vetting.  These planets
represent a superset of those in the fully automated Q1-Q17 DR25 KOI
Table \citep{Thompson_2018}, which could be adopted in future work for
planet occurrence rate calculations.  This version of the cumulative
KOI table included \nkoisnofp\ objects that are either ``confirmed''
or ``candidate'' planets, after excluding known false positives.

@@

% stat from _estimate_new_lithium_timefrac.py
Lithium equivalent widths and abundances for the Kepler Objects of
Interest were already analyzed by \citet{2018ApJ...855..115B} for
roughly three quarters of the spectra in our sample.  However, new
spectra have since been acquired, and our approach and selection
function are different.  We therefore perform our own line width
measurements on
the reduced HIRES spectra.

We collected all blaze-corrected HIRES spectra from our group's
observations of non-false positive Kepler Objects of Interest with a
``multiple event statistic'' (MES, {\it koi\_max\_mult\_ev}) of at
least 10.  This yielded at least one spectrum for \nstwithspec\
stars hosting \nplwithspec\ planets.  About half of these stars have
rotation periods (\nstwithspecandprot\ stars and
\nplwithspecandprot, respectively).   For stars with multiple spectra
available at varying ranges of signal-to-noise (S/N), we analyzed only
the spectrum with the highest number of counts.  The resulting spectra
were acquired between 6 September 2009 and 16 May 2024, and
cumulatively comprise \nhireshours\,hours of open-shutter time.

@@

Our EW measurement approach does not correct for the neighboring
\ion{Fe}{1} 6707.44\,\AA\ blend.  To evaluate the accuracy and
precision of our measurement technique, after applying an initial
iteration on the \nlithiumstars\ KOIs with spectra, we compared our
lithium equivalent widths with those reported by
\citet{2018ApJ...855..115B}.  For the \nbergeroverlap\ stars in both
samples, we found broad agreement at $>$30\,m\AA, and significant
differences at $<$30\,m\AA\ because \citet{2018ApJ...855..115B}
required positive EWs, while we allow for statistically negative ones.  At
$>$30\,m\AA, there is however a small offset in our respective scales,
such that (B24-B18)$=$7.50$\pm$8.55\,m\AA.  This offset is caused by
our non-treatment of the iron blend.  We therefore directly subtract
this constant mean value (7.50\,m\AA) when calculating lithium-based
ages.

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.47\textwidth]{li_vs_teff_koi_X_JUMP_eagles.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Equivalent widths (EW) of the \ion{Li}{1} 6708\,\AA\ doublet
    for planet-hosting stars.} These measurements were made from
    Keck/HIRES spectra collected between 2009 and 2024 that comprised
    \nhireshours\,hours of open-shutter time.  Lines are the ``mean''
    isochrones from the same models as in Figure~\ref{fig:models}.  The
    intrinsic dispersion beyond $\approx$1\,Gyr becomes much larger
    than changes in the mean.  Some stars with lithium detections do
    not have detected rotation periods, and vice-versa.
		\label{fig:li_vs_teff}
	}
\end{figure}

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{gyromodeldispersion.pdf}
		}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{li_vs_teff_koi_X_S19S21dquality_eagles_showpoints_nodata.pdf}
		}
	\end{center}
	\vspace{-0.6cm}
	\caption{
		{\bf The models.}
    Points represent $10^4$ draws from models that have been fitted to
    rotation periods \citep{Bouma_2023} and lithium equivalent widths
    \citep[EWs;][]{Jeffries_2023} of stars in open clusters.  Lines
    are the ``mean models'' at various ages.  The intrinsic dispersion
    in rotation and lithium about these mean models, which is what the
    models fit, sets the theoretical precision floor for the
    age-dating methods.    Additional sources of uncertainty,
    including measurement uncertainty, impose further limits on
    achievable precision.  These models are calibrated using open
    clusters younger than 4\,Gyr.  The displayed points assume a
    uniform distribution in temperature for visual clarity.  The sizes
    of the points are the same in each panel, so that low apparent
    density signifies greater dispersion around the mean.
		\label{fig:models}
	}
\end{figure*}

@@

Using our effective temperatures and rotation periods, we ran
\texttt{gyro-interp} over an age grid of 0 to 4\,Gyr, linearly spaced
over 500 grid points.  \texttt{gyro-interp} itself is calibrated out
to 4\,Gyr, the age of M67
\citep[see][]{2022ApJ...938..118D,Gruner_2023}.  To estimate the
stellar spin-down rate as a function of time and stellar temperature,
the framework interpolates between reference open clusters using cubic
Hermite polynomials.  We calculate the probability of the
rotation-based age $t_{\rm gyro}$, given the observed rotation
periods, effective temperatures, and their uncertainties by
integrating Equation~1 of \citet{Bouma_2023} at the default grid
resolution.  This procedure marginalizes over the intrinsic age and
temperature dependent scatter that is observed in the cluster
sequences to yield statistical age uncertainties.

@@

\subsection{Planet Quality Flags}
\label{subsec:plflags}
Some planets are more reliably identified than others.  We use the
following additional quality indicators to assess the reliability and
utility of a planet, and assemble them into a separate bit-flag,
$Q_{\rm planet}$.

@@

In the planet sample, the cumulative KOI table included
\nnonfpcumkois\ confirmed and candidate planets, orbiting
\nnonfpcumkoihosts\ stars.  Assume that we apply all the same stellar
cuts, save for that on RUWE.  We can then require these confirmed and
candidate planets to have MES$>$10.  This combination of selection
criteria yield \nplwgyroagewithgrazingandhighruwe\ planets with
rotation-based ages orbiting
\nplhoststarwgyroagewithgrazingandhighruwe\ stars
(\nplhoststarwgyroagejusthighruwe\ of these stars with ``high'' RUWE).
If we additionally require the planets to be non-grazing ($b<0.8$), we
are left with \nplwgyroagenograzing\ planets orbiting
\nplhoststarwgyroagenograzing\ stars.

@@

To calculate lithium ages, we use \texttt{EAGLES} (git commit
\texttt{ac09637}), which, similar to \texttt{gyro-interp}, is based on
a purely empirical interpolation approach.  \texttt{EAGLES} 
was calibrated on lithium measurements of stars observed by the
Gaia-ESO spectroscopic survey in 52 open clusters with ages spanning 2
to 6{,}000\,Myr \citep{Jeffries_2023}.  We adopt a linear prior in
age, spanning $10^6$ to $10^{10}$\,yr, and symmetrize our EW
uncertainties for purposes of interfacing with \texttt{EAGLES} by
taking the maximum of the measured positive and negative 1$\sigma$
uncertainty intervals.

@@

Figure~\ref{fig:hist_tgyro} shows the gyrochone age distributions of
stars and known planet-hosts in the Kepler field.  In detail, these
``histograms'' are sampled from the posterior probability
distributions by drawing ten random samples from each posterior
distribution, and then computing the histogram of the resulting
samples.  We truncated the plots near $4$\,Gyr since this is the
cluster age out to which our interpolation-based method is calibrated.

For Sun-like stars, the fraction of rotation signals detectable by
Kepler is near unity up to $\approx$3\,Gyr, and it
drops to almost zero by $\approx$5\,Gyr \citep{2022ApJ...937...94M}.
This trend has the opposite functional form from our derived age
distributions.

@@

In terms of detected planet counts, our rotation-based ages for the
Kepler sample yield \{\nplyounggyro, \nplmidgyro, \nploldgyro\}
detected planets in the 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr bins.  
The stellar ages and planetary
properties of all of these systems are given in
Table~\ref{tab:planets}.

@@

Sub-gigayear planets hold particular weight in studies of planet
evolution.
While our catalog has \nplyounggyro\ confirmed planets
with median ages below 1\,Gyr, requiring $t_{\rm gyro}$$<$1\,Gyr at
2$\sigma$ confidence yields \nplyounggyrotwosigma\ planets orbiting
\nplhostsyounggyrotwosigma\ stars.  
The most notable of these planets can be split into two classes: those
in clusters, and those with independent and consistent rotation and
lithium-based ages.

@@

Four stellar ensembles in the Kepler field have to date yielded a
total of fourteen transiting planets.  Our analysis blindly recovers
the youth of these planets.

@@

We compared the independent rotation and lithium-based ages of the
known planet hosts using the point-estimates of the available
posteriors.  There are three cases of interest:

@@

In the entire planet sample (i.e.~without imposing any quality cuts) this
yielded \allagesyesconsistent\ consistent cases,
\allagesmaybeconsistent\ ``maybe consistent'' cases, and
\allagesnoconsistent\ inconsistent cases.  Rephrased, in the overall
sample, \fracconsistentallages\% of stars have consistent rotation and
lithium-based ages (\fracpotentiallyconsistentallages\% potentially
consistent).
